Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-01 paper/tv.tex
m

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract} 
Real-time systems are systems where the correctness does not 
only depend on their correct functioning but also on meeting 
realtime constraints.
Different from existing timed-automata based techniques, 
we propose a novel solution that integrates a 
modular Hoare-style forward verifier with a new term rewriting 
system (TRS) on \emph{Timed Effects} (\timedEffects).
The main purpose is to dynamically create clocks  
and efficiently solve constraints on the clocks, 
We formally define 
a core language \timedL, generalizing the real-time systems, modeled 
using mutable variables and timed behavioral patterns, 
such as \emph{delay}, \emph{timeout}, \emph{deadline}. 
Secondly, to capture real-time specifications, 
we introduce \timedEffects, a new effects logic, 
that extends 
\emph{Regular Expressions} with dependent
values and arithmetic constraints.
Thirdly,  the forward verifier infers temporal behaviors of given 
\timedL\ programs, expressed in \timedEffects. 
Lastly, we present a purely algebraic TRS, i.e., the extended Antimirov algorithm, to 
efficiently prove language inclusions between 
 \timedEffects. 
To demonstrate the feasibility of our proposals, 
we prototype the verification system; prove its 
correctness; report on case studies and the experimental results.

@@

several compositional ,  
are introduced to capture quantitative timing constraints.
The idea is to

Specification and verification of real-time systems are important 
research topics which have practical implications. During the last 
(more than) two decades, a popular approach for specifying real-time systems is 
based on Timed Automata \cite{DBLP:journals/tcs/AlurD94}. 
Timed Automata are powerful 
in designing real-time models with explicit clock variables. Real-time 
constraints are captured by explicitly setting/reseting clock variables. 
A number of automatic verification support for Timed Automata have proven 
to be successful, such as UPPAAL \cite{DBLP:journals/sttt/LarsenPY97}.

@@

\begin{comment}
A TRS is a refutation method that normalizes expressions in such a way 
that checking their inclusion corresponds to an iterated process of 
checking the inclusion of their \emph{partial derivatives} 
\cite{antimirov1995partial}. 
Works based on such a TRS \cite{song2020automated,antimirov1995rewriting,almeida2009antimirov,keil2014symbolic,hovland2012inclusion}  show its feasibility and suggest that this method is a better average-case algorithm than those based on the comparison of automata. 
\end{comment}

@@

\begin{figure}[ht]
      \vspace{-2mm}
\begin{lstlisting}[columns=fullflexible]
hiphop module main (out Prep, in Tick, out Ready, out Go, out Cook)
/*@ requires true : emp @*/
/*@ ensures  0<=t/\t<3 : {}.({Cook})#t.{}^* @*/
{
	fork{ // The effects of the first thread is true : Ready?.{Go}
		await Ready; emit Go; 
	}par{ // The effects of the second thread is 0<t<3 : {Prep, Cook}#t.{Ready}
		emit Prep;
		async Ready { run cook (3, Tick, Cook); }}}
// The final effects for main is  0<=t/\t<3 : ({Prep}.{Cook})#t.{Ready}.{Go}

hiphop module cook (var d, in Tick, out Cook)
/*@ requires d>2  : {}^*.{Prep} @*/
/*@ ensures  0<=t/\t<d : ({}.{Cook})#t @*/
{	abort count(d, Tick) { yield; emit Cook; }}
\end{lstlisting}  
      \vspace{-1mm}
      \caption{Mixed Synchronous and Asynchronous Concurrency with Function Calls.}\label{fig:overview_eg2}
            \vspace{-1mm}
\end{figure}

%We demonstrate

%%, and a function call from module \code{main} to \code{cook}.

      \vspace{-2mm}
\subsection{Forward Verification}
As shown in \figref{fig:forward_example}, we demonstrate the forward verification process of the module  {\ttfamily{main}}. The program effects states  are captured in the form of \code{\langle \textcolor{huntergreen}{\effect} \rangle}. To facilitate the illustration, we label the verification steps by (1), ..., (9), and mark the deployed inference rules (cf. \secref{Forward_Rules}) in \textcolor{mGray}{[gray]}.

@@

To formulate the target language, we generalise the design of  Hiphop.js into a core language \code{\timedL}, 
%Based on the design of , but not too specific to Hiphop.js, we summarise a core language to be our target language, 
which provides the infrastructure for mixing synchronous and asynchronous concurrency models. 
We here formally define the syntax of \code{\timedL}, as shown in \figref{fig:Hiphop_Syntax} The statements marked as \textcolor{purple} {purple} come from the  Esterel v5 \cite{berry1999constructive,berry2esterel} endorsed by current academic compilers; while the statements marked as \textcolor{blue} {blue} provide the asynchrony coming from the usage of JavaScript promises. 
In this work, we are mainly interested in signal status and control propagation, which are not related to data, therefore the data variables and data-handling primitives are abstracted away.

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{\tau \ mn  \ {( \tau\ x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ ()\ |\ k^\tau\ |\ x\ |\ 
    %v.f\ |\ 
    %\textbf{new}\ d({v^*}) \ |\ 
    \tau\ x;\ e \ |\ mn({x^*}) \ }
  \code{|\ x{:=}e\ |\ 
 % v.f:=e \ | 
 \ e_1;e_2\ 
  |\ {if}\ v\ {then}\ e_1\ {else}\ e_2 
 }  \\
 &   & & \code{
% |\ e_1}\ op\ \code{ e_2 \ 
| \ \textbf{event} [\anyevent{A}]\
  %|\ \textbf{assert}\ \effect \ 
  | \ \textbf{timeout} \ e \ d \  
  | \ \textbf{deadline} \ e \ d \  
  | \ \textbf{delay}\ d 
  }  \\
  ~\\

    \multicolumn{4}{c}{
  \code{k^\tau:constant\ of\ type\ \tau\ } 
  \qquad\quad
\code{    x, mn  ::\in \textbf{var}}  
\qquad\quad
 \code{(Events)  \anyevent{A} ::\in \Sigma 
\qquad\quad
\m{(Time\ Bounds)}\ \code{d} \code{\in \mathbb{Z^+}}    
}   
    }\\  
    \hline
  \end{array}  
  $
 \caption{A Core Imperative Language.} 
 \label{fig:code_language}
  \vspace{-1mm}
\end{figure}
}

@@

   \code{ \bott
   \  | \ \empt 
    \ | \ \anyevent{A} 
    \ | \ \anynotevent{A}
    \ | \ \_  
    %   \ | \ \anyevent{a}
   \  | \ \es_1 {\seq} \es_2
    \ | \ \es_1 {\vee} \es_2
    \ | \ \es_1 {||} \es_2
      %    \  | \ \es_1 \times \es_2
      \ | \ \es \# t 
 %  \  | \  \esn{\es}{t}    
   \  | \  \esn{\es}{\star}    
  % \  | \ \esn{\es}{\omega} 
    }

@@

    \\~\\
    \multicolumn{4}{l}{
      \code{(Events)  \anyevent{A} ::\in \Sigma}   
   \qquad
    \code{c ::\in \mathbb{R}^{+}} 
     \qquad 
           \code{x  ::\in \textbf{var} } 
           \qquad

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract} 
Real-time systems are systems where the correctness does not 
only depend on their correct functioning but also on meeting 
realtime constraints.
Different from existing Timed Automata based techniques, 
we propose a novel solution that integrates a 
modular Hoare-style forward verifier with a new term rewriting 
system (TRS) on \emph{Timed Effects} (\timedEffects).
The main purpose is to dynamically create clocks  
and efficiently solve constraints on the clocks, 
We formally define 
a core language \timedL, generalizing the real-time systems, modeled 
using mutable variables and timed behavioral patterns, 
such as \emph{delay}, \emph{timeout}, \emph{deadline}. 
Secondly, to capture real-time specifications, 
we introduce \timedEffects, a new effects logic, 
that extends 
\emph{Regular Expressions} with dependent
values and arithmetic constraints.
Thirdly,  the forward verifier infers temporal behaviors of given 
\timedL\ programs, expressed in \timedEffects. 
Lastly, we present a purely algebraic TRS, i.e., an extended Antimirov algorithm, to 
efficiently prove language inclusions between 
 \timedEffects. 
To demonstrate the feasibility of our proposals, 
we prototype the verification system; prove its 
correctness; report on case studies and the experimental results.

@@

Specification and verification of real-time systems are important 
research topics which have practical implications. During the last 
more than two decades, a popular approach for specifying real-time systems is 
based on Timed Automata \cite{DBLP:journals/tcs/AlurD94}. 
Timed Automata are powerful 
in designing real-time models with explicit clock variables. Real-time 
constraints are captured by explicitly setting/reseting clock variables. 
A number of automatic verification support for Timed Automata have proven 
to be successful \cite{DBLP:journals/sttt/LarsenPY97,yovine1997kronos,wang2005verifying}.

@@

\begin{comment}
  several compositional ,  
are introduced to capture quantitative timing constraints.
The idea is to

A TRS is a refutation method that normalizes expressions in such a way 
that checking their inclusion corresponds to an iterated process of 
checking the inclusion of their \emph{partial derivatives} 
\cite{antimirov1995partial}. 
Works based on such a TRS \cite{song2020automated,antimirov1995rewriting,almeida2009antimirov,keil2014symbolic,hovland2012inclusion}  show its feasibility and suggest that this method is a better average-case algorithm than those based on the comparison of automata. 
\end{comment}

@@

\subsection{Forward Verification}
As shown in \figref{fig:forward_example}, we demonstrate the forward verification process of the module  {\ttfamily{main}}. The program effects states  are captured in the form of \code{\langle \textcolor{huntergreen}{\effect} \rangle}. To facilitate the illustration, we label the verification steps by (1), ..., (9), and mark the deployed inference rules (cf. \secref{Forward_Rules}) in \textcolor{mGray}{[gray]}.

@@

To formulate the target language, we generalize the design of the real-time systems
 into a core language \code{\timedL}, 
 which provides the infrastructure for mutable variables and timed behavioral patterns, 
 such as \emph{delay}, \emph{timeout}, \emph{deadline}. 
We here formally define the syntax of \code{\timedL}, as shown in \figref{fig:code_language}

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{\tau \ mn  \ {( \tau\ x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ ()\ |\ k^\tau\ |\ x\ |\ 
    %v.f\ |\ 
    %\textbf{new}\ d({v^*}) \ |\ 
    \tau\ x;\ e \ |\ mn({x^*}) \ }
  \code{|\ x{:=}e\ |\ 
 % v.f:=e \ | 
 \ e_1;e_2\ 
  |\ {if}\ v\ {then}\ e_1\ {else}\ e_2 
 }  \\
 &   & & \code{
% |\ e_1}\ op\ \code{ e_2 \ 
| \ \textbf{event} [ev]\
  %|\ \textbf{assert}\ \effect \ 
  | \ \textbf{timeout} \ e \ d \  
  | \ \textbf{deadline} \ e \ d \  
  | \ \textbf{delay}\ d 
  }  \\
  ~\\

    \multicolumn{4}{c}{
  \code{k^\tau:constant\ of\ type\ \tau\ } 
  \qquad\quad
\code{    x, mn  ::\in \textbf{var}}  
\qquad\quad
 \code{(Events)  ev ::\in \Sigma 
\qquad\quad
\m{(Time\ Bounds)}\ \code{d} \code{\in \mathbb{Z^+}}    
}   
    }\\  
    \hline
  \end{array}  
  $
 \caption{A Core Imperative Language.} 
 \label{fig:code_language}
  \vspace{-1mm}
\end{figure}
}

@@

   \code{ \bott
   \  | \ \empt 
    \ | \ ev 
    %   \ | \ \anyevent{a}
   \  | \ \es_1 {\seq} \es_2
    \ | \ \es_1 {\vee} \es_2
    \ | \ \es_1 {||} \es_2
      %    \  | \ \es_1 \times \es_2
      \ | \ \es \# t 
 %  \  | \  \esn{\es}{t}    
   \  | \  \esn{\es}{\star}    
  % \  | \ \esn{\es}{\omega} 
}

\\
\m{(Events)} & ev  & \ ::= \ & 
\code{
    \anyevent{A} (c)
    \ | \ \anynotevent{A} (c)
    \ | \ \_  
}

@@

    \\~\\
    \multicolumn{4}{c}{
     % \code{(Events)  \anyevent{A} ::\in \Sigma}   
   %\qquad
    \code{c ::\in \mathbb{R}^{+}} 
     \qquad 
           \code{x  ::\in \textbf{var} } 
           \qquad
